### 2022-10-17 labs/Chapter 17 Lab.Rmd

---
title: "Chapter 17 Lab - Random Forests and Boosting"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, include = FALSE}
library(tree) #may need to install
library(ISLR)
library(dplyr)
library(ggplot2)
library(gbm)
library(randomForest)
```

This lab is an edited version of an abbreviated version of p. 324-331 of "Introduction to Statistical Learning with Applications in R" by Gareth James, Daniela Witten, Trevor Hastie and Robert Tibshirani. It was re-implemented in Fall 2016 in `tidyverse` format by Amelia McNamara and R. Jordan Crouser at Smith College, and further edited by Amy Wagaman at Amherst College in Fall 2018, Fall 2019, and Fall 2022.

The lab reviews classification trees (as in CASI Chapter 8's classification trees)  but uses a different package to fit them called *tree*. You can use either tree or rpart. Rpart gives a bit better control, but tree has a few other options you might want. After that, it covers random forests, bagging, and boosting. The code is provided in this until the open-ended problem at the end. Be sure you are comfortable with the output, etc. Remember you can use these for regression or classification.

# Fitting Trees - Review with new R Package

The `tree` library is useful for constructing classification and regression trees:

We'll start by using **classification trees** to analyze the `Carseats` data set. In these
data, `Sales` is a continuous variable, we want to consider a binary classification problem, so we begin by converting it to a binary variable. We use the `ifelse()` function to create a variable, called `High`, which takes on a value of `Yes` if the `Sales` variable exceeds 8, and takes on a value of `No` otherwise:

```{r}
Carseats <- Carseats %>%
  mutate(High = as.factor(ifelse(Sales <= 8, "No", "Yes")))
```

In order to properly evaluate the performance of a classification tree on
the data, we must estimate the test error rather than simply computing
the training error. We first split the observations into a training set and a test
set:

```{r}
set.seed(495)

n <- nrow(Carseats)
train_index <- sample(1:n, 0.5 * n)
test_index <- setdiff(1:n, train_index)

train <- Carseats[train_index, ] 
test <- Carseats[test_index, ]
```

We now use the `tree()` function to fit a classification tree in order to predict
`High` using all variables but `Sales` (that would be a little silly...). The syntax of the `tree()` function is quite similar to that of the `lm()` function:

```{r}
tree_carseats <- tree(High ~ . - Sales, train)
```

The `summary()` function lists the variables that are used as internal nodes (forming decision points)
in the tree, the number of terminal nodes, and the (training) error rate:

```{r}
summary(tree_carseats)
```

One of the most attractive properties of trees is that they can be
graphically displayed. We use the `plot()` function to display the tree structure,
and the `text()` function to display the node labels. The argument
`pretty = 0` instructs `R` to include the category names for any qualitative predictors,
rather than simply displaying a letter for each category:

```{r}
plot(tree_carseats)
text(tree_carseats, pretty = 0)
```

Again, you've seen nicer plots are possible with rpart.plot for rpart objects.

The most important indicator of `High` sales appears to be shelving location,
since the first branch differentiates `Good` locations from `Bad` and `Medium`
locations.

If we just type the name of the tree object, `R` prints output corresponding
to each branch of the tree. `R` displays the split criterion (e.g. $Price < 142$), the
number of observations in that branch, the deviance, the overall prediction
for the branch (`Yes` or `No`), and the fraction of observations in that branch
that take on values of `Yes` and `No`. Branches that lead to terminal nodes are
indicated using asterisks:

```{r}
tree_carseats
```

Finally, let's evaluate the tree's performance on
the test data. The `predict()` function can be used for this purpose. In the
case of a classification tree, the argument `type="class"` instructs `R` to return
the actual class prediction.

```{r}
tree_pred <- predict(tree_carseats, test, type = "class")
table(tree_pred, test$High)
```

# Pruning

Next, we consider whether pruning the tree might lead to improved
results. The function `cv.tree()` performs cross-validation in order to
determine the optimal level of tree complexity; cost complexity pruning
is used in order to select a sequence of trees for consideration. We use
the argument `FUN = prune.misclass` in order to indicate that we want the
classification error rate as our cost function to guide the cross-validation and pruning process,
rather than the default for the `cv.tree()` function, which is `deviance`. The
`cv.tree()` function reports the number of terminal nodes of each tree considered
(size) as well as the corresponding error rate and the value of the
cost-complexity parameter used.

```{r}
set.seed(495)
cv_carseats <- cv.tree(tree_carseats, FUN = prune.misclass)
```

Note that, despite the name, the `dev` field corresponds to the cross-validation error
rate in this instance. Let's plot the error rate as a function of size:

```{r}
plot(cv_carseats$size, cv_carseats$dev, type = "b")
```

We see from this plot that the tree with 12 terminal nodes results in the lowest
cross-validation error rate.

We now apply the `prune.misclass()` function in order to prune the tree to
obtain the ten-node tree by setting the parameter `best = 12`:

```{r}
prune_carseats <- prune.misclass(tree_carseats, best = 12)
plot(prune_carseats)
text(prune_carseats, pretty = 0)
```

How well does this pruned tree perform on the test data set? Once again,
we can apply the `predict()` function top find out:

```{r}
tree_pred <- predict(prune_carseats, test, type = "class")
table(tree_pred, test$High)
```

We can check to see what percent of the test observations are correctly classified, and use that to determine if  
the pruning process resulted in a large or small change in classification accuracy. There can be a tradeoff between an interpretable tree versus prediction accuracy.

Note that setting different seeds can lead to different trees and different CV pruning results. This is part of the instability of trees on different data sets.

# Regression Trees

Now let's try fitting a **regression tree** to the `Boston` data set from the `MASS` library. First, we create a
training set, and fit the tree to the training data using `medv` (median home value) as our response:

```{r}
set.seed(495)
n <- nrow(MASS::Boston)
train_index <- sample(1:n, 0.5 * n)
test_index <- setdiff(1:n, train_index)

boston_train <- MASS::Boston[train_index, ] 
boston_test <- MASS::Boston[test_index, ]

tree_boston <- tree(medv ~ ., boston_train)

summary(tree_boston)
```

Notice that the output of `summary()` indicates that only a few of the variables
have been used in constructing the tree. In the context of a regression
tree, the `deviance` is simply the sum of squared errors for the tree. Let's
plot the tree:

```{r}
plot(tree_boston)
text(tree_boston, pretty = 0)
```

The variable `lstat` measures the percentage of individuals with lower
socioeconomic status. The tree indicates that lower values of `lstat` correspond
to more expensive houses.

Now we use the `cv.tree()` function to see whether pruning the tree will
improve performance:

```{r}
cv_boston <- cv.tree(tree_boston)
plot(cv_boston$size, cv_boston$dev, type = 'b')
```

It looks like the full tree (or near to it) has the lowest error, but to practice with pruning, we can see the plot levels off at around size 7.

So, let's prune to the 7-node tree. We can prune the tree using the
`prune.tree()` function as before:

```{r}
prune_boston <- prune.tree(tree_boston, best = 7)
plot(prune_boston)
text(prune_boston, pretty = 0)
```

Now we'll use the pruned tree to make predictions on the test set:

```{r}
single_tree_estimate <- predict(prune_boston, 
                               newdata = boston_test)
ggplot() + 
    geom_point(aes(x = boston_test$medv, y = single_tree_estimate)) +
    geom_abline()

mean((single_tree_estimate - boston_test$medv)^2)
```

In other words, the test set MSE associated with the regression tree is
21.127. The square root of the MSE is therefore around 4.596, indicating
that this model leads to test predictions that are within around $4,596 of
the true median home value for the suburb.

# Bagging and Random Forests

Let's see if we can improve on this result using bagging and random forests. Recall that bagging is simply a special case of a random forest with $m = p$. The `randomForest()` function can
be used to perform both random forests and bagging. By default, 500 trees are "grown". You can change this and other options. Let's start with bagging:

```{r}
set.seed(495)
bag_boston <- randomForest(medv ~ ., 
                          data = boston_train, 
                          mtry = 13, 
                          importance = TRUE)
bag_boston
```

The argument `mtry = 13` indicates that all 13 predictors should be considered
for each split of the tree -- in other words, that bagging should be done. How
well does this bagged model perform on the test set?

```{r}
bagged_estimate <- predict(bag_boston, 
                          newdata = boston_test)

ggplot() + 
    geom_point(aes(x = boston_test$medv, y = bagged_estimate)) +
    geom_abline()

mean((bagged_estimate - boston_test$medv)^2)
```

The test set MSE associated with the bagged regression tree is dramatically smaller than that obtained using an optimally-pruned single tree! We can change
the number of trees grown by `randomForest()` using the `ntree` argument:

```{r}
bag_boston_25_trees <- randomForest(medv ~ ., data =  boston_train, mtry = 13, ntree = 25)
bagged_estimate_25_trees <- predict(bag_boston_25_trees, newdata = boston_test)
mean((bagged_estimate_25_trees - boston_test$medv)^2)
```

We can grow a random forest in exactly the same way, except that
we'll use a smaller value of the `mtry` argument. By default, `randomForest()`
uses $p/3$ variables when building a random forest of regression trees, and
$\sqrt{p}$ variables when building a random forest of classification trees. Here we'll
use `mtry = 6`:

```{r}
set.seed(495)
rf_boston <- randomForest(medv ~ ., 
                         data = boston_train, 
                         mtry = 6, 
                         importance = TRUE)

random_forest_estimate <- predict(rf_boston, 
                                 newdata = boston_test)

mean((random_forest_estimate - boston_test$medv)^2)
```

The test set MSE is quite low, although bagging seems slightly better in this case (at least for the seed I set).

Using the `importance()` function, we can view the importance of each variable:

```{r}
importance(rf_boston)
```

Two measures of variable importance are reported. The former is based
upon the mean decrease of accuracy in predictions on the out-of-bag samples
when a given variable is excluded from the model. The latter is a measure
of the total decrease in node impurity that results from splits over that
variable, averaged over all tree. In the case of regression trees, the node impurity is measured by the training RSS, and for classification trees by the deviance. Plots of these importance
measures can be produced using the `varImpPlot()` function:

```{r}
varImpPlot(rf_boston)
```

The results indicate that across all of the trees considered in the random
forest, the wealth level of the community (`lstat`) and the house size (`rm`)
are by far the two most important variables.

# Boosting

Now we'll use the `gbm` package, and within it the `gbm()` function, to fit boosted
regression trees to the `Boston` data set. We run `gbm()` with the option
`distribution = "gaussian"` since this is a regression problem; if it were a binary
classification problem, we would use `distribution = "bernoulli"`. The
argument `n.trees = 5000` indicates that we want 5000 trees, and the option
`interaction.depth = 4`limits the depth of each tree. A depth of 4 is actually pretty common here, you don't want to go much bigger. Using a depth of 1 is just stumps.

```{r}
set.seed(495)
boost_boston <- gbm(medv ~ ., 
                   data = boston_train, 
                   distribution = "gaussian", 
                   n.trees = 5000, 
                   interaction.depth = 4)
```

The `summary()` function produces a relative influence plot and also outputs
the relative influence statistics:

```{r}
summary(boost_boston)
```

We see that `lstat` and `rm` are again the most important variables by far. We can
also produce partial dependence plots for these two variables. These plots
illustrate the marginal effect of the selected variables on the response after
integrating out the other variables. In this case, as we might expect, median
house prices are increasing with `rm` and decreasing with `lstat`:

```{r}
plot(boost_boston, i = "rm")
plot(boost_boston, i = "lstat")
```

Now let's use the boosted model to predict `medv` on the test set:

```{r}
boost_estimate <- predict(boost_boston, 
                         newdata = boston_test, 
                         n.trees = 5000)

mean((boost_estimate - boston_test$medv)^2)
```

The test MSE obtained is higher than the test MSE for random forests and bagging. If we want to, we can perform boosting with a different value of the shrinkage parameter $\lambda$, which is $\epsilon$ in the CASI textbook. The default value is 0.001, but this is easily modified. Here we take $\lambda = 0.1$:

```{r}
boost_boston2 <- gbm(medv ~ ., data = boston_train, 
                    distribution = "gaussian", 
                    n.trees = 5000, 
                    interaction.depth = 4, 
                    shrinkage = 0.01, 
                    verbose = F)

boost_estimate2 <- predict(boost_boston2, newdata = boston_test, n.trees = 5000)
mean((boost_estimate2-boston_test$medv)^2)
```

In this case, using $\lambda = 0.1$ leads to a slightly lower test MSE than $\lambda = 0.001$.

# Practice

We use a data set that we previously examined when learning about ridge regression that we are returning to for practice.

```{r}
data(Hitters)
# ?Hitters # MLB data from 1986 and 1987; 322 obs on 20 variables
```

We want to predict Salary, which is missing for 59 players.

```{r}
Hitters <- na.omit(Hitters) #263 obs now
```

Use bagging, random forests, and boosting to explore models for predicting salary. Be sure to explore model options and the various parameter settings. (You can fit a single tree for comparison as well, if you want.)

> SOLUTION

```{r}

```