### 2022-10-20 homework/hmk5/hmk5Stat495F22.Rmd
Finished Hmk5 and created pdf

---
title: "Homework 5 - Stat 495"
author: "Justin Papagelis"
date: "Due Wednesday, October 19th by midnight"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, include = FALSE}
library(mosaic)
options(digits = 6)
library(rpart)
library(partykit)
library(GGally)
library(broom)
library(ggplot2)
library(lmtest) # for likelihood ratio tests

require(gridExtra) 
library(rpart.plot)
library(ISLR)
library(glmnet)

library(lars)    # you may need to install some of these (one time only)
library(leaps)   # to use variable selection methods if desired
```

# Practicing Academic Integrity

If you worked with others or used resources outside of provided course material (anything besides our textbook(s), course materials in Moodle/Git repo, R help menu) to complete this assignment, please acknowledge them below using a bulleted list.

<!-- ~~~~~~~~~~~~~~~~ YOU MAY BEGIN EDITING BELOW THIS LINE ~~~~~~~~~~~~~~~~ -->

*I acknowledge the following individuals with whom I worked on this assignment:*

Name(s) and corresponding problem(s)

*

*I used the following sources to help complete this assignment:*

Source(s) and corresponding problem(s)

* https://www.kaggle.com/datasets/swathiachath/kc-housesales-data
* https://statisticsglobe.com/count-number-of-na-values-in-vector-and-column-in-r
* https://cran.r-project.org/web/packages/rpart/vignettes/longintro.pdf

\newpage

# Introduction and Exploratory Data Analysis 
We are interested in determining the features predictive of homes selling for more than half a million dollars. We are only interested in whether the price is over or under \$500,000, so we will transform it into a binary response variable where the price is either over \$500,000 or is not. Then because the response variable is binary, we will use logistic regression to create the best model we can. We will also perform a non-parametric analysis using classification trees to help visualize what predictors are best indicators of a house selling for greater than \$500,000. Finally, we can compare the models to help answer the real estate developer's questions.

We used the King County, Washington (State) house sales data set (`kchouse`) for this analysis. Reference for data: https://www.kaggle.com/swathiachath/kc-housesales-data/version/1. There is no information on the source of the data.

```{r}
kchouse <- read.csv("https://awagaman.people.amherst.edu/stat495/kc_house_data.csv", header = T)
```

The data set contains the following variables:
\begin{itemize}
\item id - notation for a specific house - Numeric
\item date - date the house was sold - String
\item price - price the house sold for in dollars - Numeric
\item bedrooms - number of bedrooms in house - Numeric
\item bathrooms- number of bathrooms per bedroom - Numeric
\item sqft\textunderscore living - square footage of the home - Numeric
\item sqft\textunderscore lot - square footage of the lot - Numeric
\item floors - total floors (levels) in house - Numeric
\item waterfront - house has a waterfront view - Y/N
\item view - house has been viewed - Numeric
\item condition - How good the condition is (overall) - Numeric
\item grade - overall grade given to housing unit (based on King County grading system: 1 poor, 13 excellent) - Numeric
\item sqft\textunderscore above - square footage of house apart from basement - Numeric
\item yr\textunderscore built - the year the house was built - Numeric
\item yr\textunderscore renovated - the year when the house was generated
\item zipcode - the zipcode the house is in - Numeric
\item lat - Latitude coordinate of house - Numeric
\item long - Longitude coordinate of house - Numeric
\item sqft\textunderscore living15 - living room area in 2015 (may or may not have been affected the lot size area) - Numeric
\item sqft\textunderscore lot15 - lot size area in 2015 - Numeric
\end{itemize}

Before we fit any models, we need to adjust how R treats the data. For example, we need to change the `date` to a date format and treat `waterfront` and `zipcode` as categorical variables instead of numeric variables. We also changed the missing values from `yr_renovated` and `sqft_basement` to "na" values instead of 0.

```{r}
# change how variables are treated by R
kchouse <- kchouse %>%
  mutate(date = lubridate::mdy(date),
         waterfront = as.factor(waterfront),
         zipcode = as.numeric(as.factor(zipcode)))
# change the values that are 0 to na which is more appropriate
kchouse <- kchouse %>%
  mutate_at(c('yr_renovated', 'sqft_basement'), ~na_if(.,0))
```

After examining `yr_renovated` and `sqft_basement`, it appears that 95.8% and 60.7% of the values, respectively, are missing. From Figure 1, there are weak relationships between `yr_renovated` and `price` as well as `sqft_basement` and `price`. Therefore, we will drop these predictors when creating our models because they will not be as helpful as other predictors.

```{r}
# calculate the percentage of missing observations in yr_renovated
sum(is.na(kchouse$yr_renovated))/nrow(kchouse)

# calculate the percentage of missing observations in sqft_basement
sum(is.na(kchouse$sqft_basement))/nrow(kchouse)
```

```{r, warning = FALSE}
# generate plot of yr_renovated and price
fig1 <- ggplot(data = kchouse, aes(x = yr_renovated, y = price)) + 
  geom_point() + 
  labs(title = "Scatterplot of Year Renovated and Price", x = "Year Renovated",
       y = "Price")

# generate plot of ysqft_basement and price
fig2 <- ggplot(data = kchouse, aes(x = sqft_basement, y = price)) + 
  geom_point() + 
  labs(title = "Scatterplot of Sq Ft Basement and Price", 
       x = "Square Feet Basement", y = "Price")
```

```{r, include = TRUE, fig.width=9, fig.height=4, warning=FALSE, fig.cap="Scatter plots showing relationship between yr_renovated and price as well as sqft_basement and price."}
grid.arrange(fig1, fig2, ncol = 2)
```

Additionally, from Figure 2, there appears to be an outlier (with 33 bedrooms) so we will remove that observation and show the relationship once the outlier is removed.

```{r, fig.cap="scatter plot between bedrooms and price. An outlier is visible."}
# generate plot of bedrooms and price
fig3 <- ggplot(data = kchouse, aes(x = bedrooms, y = price)) + 
  geom_point() + 
  labs(title = "Scatterplot of Bedrooms and Price", 
       x = "Bedrooms", y = "Price")

# remove observation
kchouse1 <- kchouse %>%
  filter(id != 2402100895)

# generate plot of bedrooms and price w/o outlier
fig4 <- ggplot(data = kchouse1, aes(x = bedrooms, y = price)) + 
  geom_point() + 
  labs(title = "Scatterplot of Bedrooms and Price (no outlier)", 
       x = "Bedrooms", y = "Price")
```

```{r, include = TRUE, fig.width=9, fig.height=4, message=FALSE, fig.cap="Scatter plots showing Bedrooms and Price before and after the removal of an outlier"}
grid.arrange(fig3, fig4, ncol = 2)
```

Since a couple of variables, such as `date` and `id` do not have a relationship with `price` as seen in Figure 3, we will remove those predictors as well.

```{r}
fig5 <- ggplot(data = kchouse, aes(x = date, y = price)) + 
  geom_point() + 
  labs(title = "Scatterplot of Date and Price", 
       x = "Date", y = "Price")

fig6 <- ggplot(data = kchouse, aes(x = id, y = price)) + 
  geom_point() + 
  labs(title = "Scatterplot of ID and Price", 
       x = "ID", y = "Price")
```

```{r, include = TRUE, fig.width=9, fig.height=4, message=FALSE, fig.cap="Scatter plots showing there is no relationship between date and price as well as ID and price."}
grid.arrange(fig5, fig6, ncol = 2)
```

Additionally, there are concerns with multicollinearity between `sqft_living` and other variables, including `sqft_living15` and `sqft_above` (Figure 4), so we are going to remove `sqft_living` as well.

```{r}
fig7 <- ggplot(data = kchouse, aes(x = sqft_living, y = sqft_above)) + 
  geom_point() + 
  labs(title = "Scatterplot of SqFt Living and SqFt Above", 
       x = "Sq Ft Living", y = "SqFt Above")

fig8 <- ggplot(data = kchouse, aes(x = sqft_living, y = sqft_living15)) + 
  geom_point() + 
  labs(title = "Scatterplot of SqFt Living and SqFt Living 2015", 
       x = "Sq Ft Living", y = "SqFt Living in 2015")
```

```{r, include = TRUE, fig.width=9, fig.height=4, message=FALSE, fig.cap="scatter plots showing multicollinearity concerns with sqft_living"}
grid.arrange(fig7, fig8, ncol = 2)
```

```{r}
kchouse1 <- kchouse1 %>%
  select(-c(id,zipcode, yr_renovated, date, sqft_basement, sqft_living))
```

Referencing Figure 5, the houses are bunched near each other (close to Seattle), so the results of this analysis can only be generalized to that region of the state.

```{r, fig.cap="Scatter plot of latitude and longitude. Clearly the houses are not dispersed."}
ggplot(data = kchouse, aes(x = lat, y = long)) +
  geom_point() + 
  labs(title = "Scatterplot of Latitude and Longitude", 
       x = "Latitude", y = "Longitude")
```

```{r, fig.cap="Histogram of Price"}
gf_dhistogram(~ price, data = kchouse) %>%
  gf_labs(title = "Histogram of Price", x = "Price")

favstats(~ price, data = kchouse)
```

From Figure 6, the distribution of `price` is right-skewed with a median of \$450,000 and an IQR is \$323,000. There appear to be many outliers on the high end for `price`. The minimum `price` is \$78,000, and the maximum is \$7,700,000. Since the distribution is right-skewed, a log transformation could be helpful if we were predicting price.

Next, since we want to know if the price of the house is greater or less than \$500,000, we will transform the price variable so that it is a binary predictor that is `1` if the price of the house is greater than \$500,000 and `0` if the price of the house is less than \$500,000. There are 12,546 houses with a price below \$500,000 and 9,050 houses with a price above \$500,000.

```{r}
kchouse1 <- mutate(kchouse1, price = ifelse(price > 500000, 1, 0))
tally(~price, data = kchouse1)
```

The last thing we will do before creating the models is creating a training and test set. We use the training set to create a model, and then the test set is used to see how the model performs in making predictions. We will randomly create a 75/25 split for the training/test set.

```{r}
# create the training and test set for price
set.seed(1)
n <- nrow(kchouse1)
train_index <- sample(1:n, 3/4 * n)
test_index <- setdiff(1:n, train_index)

train <- kchouse1[train_index, ]
test <- kchouse1[test_index, ]
```

\newpage

# GLM and Results

The first models we will be fitting are GLMs (or Generalized Linear Models), extensions of ordinary linear regression (fitting a least-squares curve) to cases where the response variables are in an exponential family form. These cases can be response variables that are binomial, Poisson, gamma, or beta. Since we transformed `price` into a binary variable, it can now be modeled as a binomial distribution. The model for a binary response variable is a logistic regression model.

First, we will fit the logistic model using all of the variables we have not removed from the data set. We use the training set to create the model.

```{r}
logmod <- glm(price ~ ., data = train, family = "binomial")
msummary(logmod)
```

The model shows that we have 9 significant predictors, and we can perform a likelihood ratio test to determine if the overall model is significant.

```{r}
lrtest(logmod)
```

Since we have a significant p-value, the overall model appears significant. Now, to make predictions using this model, we need to bin the fitted values, which we will do below.

```{r}
augprice1 <- augment(logmod, type.predict = "response")
augprice1 <- mutate(augprice1, binprediction = round(.fitted, 0))
tally(~ binprediction, data = augprice1)
9826/(9826+6371)

```

After using 0.5 as a default cutoff, the model classifies 60.7% of the prices as below \$500,000. This is close to the actual percentage of the training data that is classified as below \$500,000, which is 58.1% of the observations. We can try to match this better by adjusting the cutoff to which we round the numbers. Finding the appropriate quantile, we now see that the percentage of the binned fitted values for prices below \$500,000 is 58.4%, much closer than before.

```{r}
# find out the actual percentage in training set
tally(~ price, data = augprice1)
9415/(9415+6782)

# find the appropriate quantile
with(augprice1, quantile(.fitted, 0.58128))

# re-bin the values and find out the percentage
augprice1 <- mutate(augprice1, binprediction2 = as.numeric(.fitted > 0.454465))
tally(~ binprediction2, data = augprice1)
9455/(9455+6742)
```

Next, we can create a confusion matrix and determine how many predictions were correct. From the matrix and calculation below, it appears that 84.0% of the predictions for the training set were correct.

```{r}
with(augprice1, table(price, binprediction2))
(8138+5465)/(8138+5465+1277+1317)
```

We will do the same on the test set to see how the model performs. We follow the same process as above to calculate the predicted values and then bin them using our previous cutoff. The model classifies 58.9% of the prices below \$500,000.

```{r}
augprice2 <- mutate(test, fitted = predict(logmod, newdata = test, type = "response"))
augprice2 <- mutate(augprice2, binprediction = as.numeric(fitted > 0.454465))
tally(~ binprediction, data = augprice2)
(3181)/(3181+2218)
```

Next, we generate a confusion matrix, and 83.8% of the predictions were correct.

```{r}
with(augprice2, table(price, binprediction))
(2719+1806)/(2719+1806+462+412)
```

Next, we will try creating a smaller model by removing some of the predictors that were not significant in our previous model. We will remove `bedrooms`, `long`, and `sqft_lot15`. After creating this model, we then repeated the process. We removed other variables that were significant but did not seem to have much of an impact on the predictions because the other variables were much more influential. The other variables we removed were `sqft_lot`, `floors`, `waterfront`, `view`, `condition`, and  `sqft_above`.

```{r}
logmod2 <- glm(price ~ bathrooms + grade + yr_built + lat + sqft_living15, 
  data = train, family = "binomial")
msummary(logmod2)
lrtest(logmod2)
```

All of the variables in our reduced model are significant, and the overall model is significant. Now, we can create some predictions following the same process as above.

```{r}
augprice3 <- augment(logmod2, type.predict = "response")
augprice3 <- mutate(augprice3, binprediction = round(.fitted, 0))
tally(~ binprediction, data = augprice3)
9829/(9829+6386)

```

After using 0.5 as a default cutoff, the model classifies 60.6% of the prices as below \$500,000. After adjusting the quantile, we now see that the percentage of the binned fitted values for prices below \$500,000 is 58.1%, much closer than before.

```{r}
# find the appropriate quantile
with(augprice3, quantile(.fitted, 0.58128))

# re-bin the values and find out the percentage
augprice3 <- mutate(augprice3, binprediction2 = as.numeric(.fitted > 0.451563))
tally(~ binprediction2, data = augprice3)
9414/(9414+6783)
```

Next, we can create a confusion matrix and determine how many predictions were correct. From the matrix and calculation below, it appears that 84.0% of the predictions for the training set were correct.

```{r}
with(augprice3, table(price, binprediction2))
(8119+5486)/(8119+5486+1296+1296)
```

We will do the same on the test set to see how the model performs. We follow the same process as above to calculate the predicted values and then bin them using our previous cutoff. The model classifies 58.6% of the prices below \$500,000.

```{r}
augprice4 <- mutate(test, fitted = predict(logmod2, newdata = test, type = "response"))
augprice4 <- mutate(augprice4, binprediction3 = as.numeric(fitted > 0.451563))
tally(~ binprediction3, data = augprice4)
(3164)/(3164+2235)
```

Next, we generate a confusion matrix, and it appears that 83.1% of the predictions were correct.

```{r}
with(augprice4, table(price, binprediction3))
(2690+1794)/(2690+1794+474+441)
```

Overall, We created two logistic regression models that performed very similarly, with the second one having a slightly lower correct prediction rate. However, for the second model, we removed many predictors that were not influential to the model and therefore were not helping make the predictions. After those variables were removed, the logistic model became simpler with fewer predictors. The real estate developer's question of interest is to find the variables indicative of houses that sell for over \$500,000. We can find the most important predictors by keeping a similar performance of the model but removing extraneous predictors. The second model contained the predictors: `bathrooms`, `grade`, `yr_built`, `lat`, and `sqft_living15`.

\newpage

# Tree and Results
The second method that we are going to be using to fit models is classification trees. Classification trees are a nonparametric method of prediction and are easy to interpret. Classification trees are used when the response variable is binary, so it is appropriate to use them to predict if the price is over or under \$500,000. They are created through recursive partitioning to create "splits" between observations with similar qualities.

We will first create a tree with the default stopping criteria.

```{r, fig.width = 6, fig.height = 5, fig.cap= "Classification Tree 1"}
price.rpart1 <- rpart(price ~ ., data = train, method = "class")
printcp(price.rpart1)
rpart.plot(price.rpart1)
```

The classification tree above uses the predictors, `grade`, `lat`, `sqft_above`, `sqft_living15`, `sqft_lot15`, and `yr_built`. The error from this classification tree is 0.8589, which is calculated from the root node error and the relative error in the output above.

```{r}
1 - 0.3368*0.4187
```

Next, we will use the classification tree we created to see how well it predicts the training set. From the confusion matrix below, the classification tree correctly predicts 85.9% of the observations.

```{r}
train1 <- mutate(train, predprice = predict(price.rpart1, type = "class"))
tally(predprice ~ price, data = train1)
(8227+5696)/(8227+5696+1188+1096)

```

Furthermore, doing the same thing for the test set, it appears that the model predicts correctly 85.5% of the observations.

```{r}
test1 <- mutate(test, predprice = predict(price.rpart1, type = "class", newdata = test))
tally(predprice ~ price, data = test1)
(2748+1870)/(2748+1870+398+383)
```

Before we create the second model, we examine the cross-validation process that was used in creating the previous model. The default cross-validation is 10-fold cross-validation. Below we can see that as the complexity parameter reaches 0.16, the relative error calculated from the cross-validation decreases slightly as the complexity parameter increases. Therefore, our next tree will use a complexity parameter of 0.016.

```{r, fig.cap="Graph of Relative Error from Cross-Validation as Complexity Parameter (cp) increses"}
plotcp(price.rpart1)
```

We are also going to control `minsplit = 500`, which means that there needs to be 500 observations in a node before a split is attempted. This is reasonable for a tree created from our training data with a size of 16197. We also set `minbucket = 300`, which means that the minimum number of observations in any terminal node is 300 observations. These controls are to try and attempt to build a smaller tree while keeping the same effectiveness of the first model.

```{r}
price.control <- rpart.control(minsplit = 500, minbucket = 300, cp = 0.016)
price.rpart2 <- rpart(price ~ ., data = train, method = "class", control = price.control)
printcp(price.rpart2)
```

The classification tree above uses the predictors `grade`, `lat`, and `sqft_living15`. The error from this classification tree is 0.8303, which is calculated from the root node error and the relative error in the output above.

```{r}
1 - 0.4053*0.4187
```

From the output below, the classification tree we created is much simpler.

```{r, fig.cap= "Classification Tree 2"}
rpart.plot(price.rpart2)
```

We will now see how the model performs in classifying observations.

```{r}
train2 <- mutate(train, predprice = predict(price.rpart2, type = "class"))
tally(predprice ~ price, data = train2)
(8054+5394)/(8054+5394+1361+1388)
```

It appears the model classifies observations 83.0% correctly in the training set. We will do the same for the test set.

```{r}
test2 <- mutate(test, predprice = predict(price.rpart2, type = "class", newdata = test))
tally(predprice ~ price, data = test2)
(2710+1772)/(2710+1772+421+496)
```

It appears that the model classifies observations 83.0% correctly in the test set as well.

The first one is more complex from these two models and correctly predicts about 2.5-3% better than the second model. However, this means that the first classification tree is less interpretable than the second one. The first tree has 11 splits, whereas the second one only has 5. Because of this reason, we would choose the second model over the first one. The slight decrease in correct predictions matters less than the increase in interpretability for the model. Additionally, the error of the second classification tree is slightly less than that of the first tree. Also, since we are determining which variables are the best predictors if a house is above \$500,000, the second model only uses 3 predictors, whereas the first model uses 6. The predictors used in the second model are `grade`, `lat`, and `sqft_living15`.

\newpage

# Conclusion
This analysis aimed to determine the variables that are most indicative of a house's price being greater than \$500,000. To accomplish this, we first created two logistic regression models. We decided that the one with fewer predictors was better due to its comparable prediction accuracy and simplicity. This model contained the predictors `bathrooms`, `grade`, `yr_built`, `lat`, and `sqft_living15`, which means that for the logistic regression model, these variables contributed the most to determine if a house had a price greater than half a million dollars. Secondly, we created two classification trees which we used to predict if a price was greater than \$500,000 or not. Again, we chose the model with a similar prediction accuracy and fewer predictors. This tree was also much smaller and thus easier to interpret. The predictors used in this model are `grade`, `lat`, and `sqft_living15`. The classification tree models performed slightly better than the logistic regression trees, but they were all within a small range so they are comparable. Combining information from both of our final models, `grade`, `lat`, and `sqft_living15` appeared as significant predictors in both models, so these variables, especially, are the ones that are most predictive of a house with a price over \$500,000.