### 2022-12-08 project/PaperTemplate/paper.tex

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf A Comparision of Bootstrap Confidence Interval Methods}

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge '\,'
Profossor Wagaman} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf A Comparision of Bootstrap Confidence Interval Methods}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
This paper explores different bootstrap methods for creating confidence
intervals and tests their performance against each other for different
types of distributions. This paper will re-introduce the ideas of
bootstraps and confidence intervals, as well as the theory behind
different methods of bootstrapped confidence intervals. Along with the
theoretical details of each method, we will provide an example of each
bootstrapped confidence interval method with a toy dataset. Then, we
will run a simulation that evaluates the performance of a couple of the
methods we introduced on a variety of sample sizes and types of
distributions.
\end{abstract}

\noindent%
{\it Keywords:} simulation, percentile, bias-corrected, acceleration, studentized
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Bootstrapping is an essential tool in Statistics, even more so now that
there is access to higher computing power. We use bootstrapping to
estimate the desired population parameter from a given sample without
making assumptions about any underlying distributions of the sample.
Different bootstrap techniques are developed, but we will focus on the
non-parametric bootstrap for our purposes. One way to make a statistical
inference is through confidence intervals which give a range of
estimates for the unknown population parameter at a certain confidence
level. We can use bootstrapping to create accurate approximate
confidence intervals for our population parameter even without knowing
its underlying distributions. There are many different ways to create
intervals, and we will go through a couple of them: The Standard Method,
The Percentile Method, The Bias-Corrected (BC) Method, The
Bias-Corrected with Acceleration (BCa) Method, and finally, The
Studentized Method. Additionally, we will go through an example of
creating a bootstrapped confidence interval for each method. Then, we
will perform a simulation to evaluate the performance of some of the
bootstrapped confidence interval methods on different types of
distributions.

\hypertarget{exposition}{%
\section{Exposition}\label{exposition}}

\hypertarget{the-non-parametric-bootstrap}{%
\subsection{The Non-Parametric
Bootstrap}\label{the-non-parametric-bootstrap}}

Bootstrapping is a statistical method of resampling that allows the
estimation of a test statistic from an unknown distribution. In
particular, bootstrapping is a computational heavy method which is
useful for many different situations.

First, we introduce the non-parametric bootstrap. Suppose we have a
random sample \(X = (x_1,x_2,\dots,x_n)\) from our unknown distribution,
\(F\) and a statistic of interest, \(\hat{\theta} = \hat{\theta}(X)\)
\citep{EfronCasi}. Ideally, the desired test statistic could be found by
repeatedly sampling new reproductions of \(X\) from \(F\). However \(F\)
is unknown, so this is not possible. The non-parametric bootstrap
creates an estimate \(\hat{F}\) from \(F\) using our sample, \(X\),
without making any parametric assumptions about \(F\) (such as its
distribution type). Therefore, the bootstrap sample could be represented
as \(X^* = (x^*_1, x^*_2, \dots, x^*_n)\) where each \(x^*_i\) is
sampled randomly with equal probability and with replacement from our
original sample: \(\{x_1,x_2,\dots,x_n\}\). From this bootstrap sample,
a bootstrap replication of the test statistic can be computed using
\(\hat{\theta}^* = \hat{\theta}(X^*)\). A large number, \(B\), of
bootstrap samples are drawn independently and the corresponding
bootstrap replication of the test statistic is calculated.
\[\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B.\]
The bootstrap estimate of the test statistic is the empirical value of
the test statistic from all of the \(\hat{\theta}^{*b}\) replications.
As \(B\) increases, \(\hat{F}\) approaches \(F\) which means that the
test statistic of interest approaches its true value as well.

\hypertarget{using-the-non-parametric-bootstrap}{%
\subsubsection{Using the Non-Parametric
Bootstrap}\label{using-the-non-parametric-bootstrap}}

We demonstrate performing a non-parametric bootstrap below using
\texttt{SnowGR} which gives the official snowfall dataset by month for
Grand Rapids, MI, starting in 1893. We will be using the \texttt{Dec}
variable which is the number of inches of snow that fell in December of
each year. In particular, we are interested in the sample mean and later
finding confidence intervals for population mean of \texttt{Dec}. The
distribution is below (Figure 1).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{data}\NormalTok{(SnowGR)}
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 1: Histogram of SnowGR"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/SnowGR histogram-1.pdf} Next,
we will perform the non-parametric bootstrap using 10,000 replications.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{orig\_mean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\CommentTok{\# theta hat}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{nboot }\OtherTok{\textless{}{-}} \DecValTok{10000}

\NormalTok{mean }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nboot)}
\NormalTok{sd }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nboot)}
\NormalTok{se }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nboot)}
\NormalTok{t\_stat }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nboot)}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{nboot) \{}
\NormalTok{  resampled }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(mosaic}\SpecialCharTok{::}\FunctionTok{resample}\NormalTok{(SnowGR, }\AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{))}
\NormalTok{  mean[i] }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{Dec, }\AttributeTok{data =}\NormalTok{ resampled)}
\NormalTok{  sd[i] }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{Dec, }\AttributeTok{data =}\NormalTok{ resampled)}
\NormalTok{\}}

\NormalTok{dec\_means }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(mean)}
\end{Highlighting}
\end{Shaded}

Below, in Figure 2, is the histogram of bootstrapped values. Since we
performed a large number of bootstrap replications, the bootstrapped
distribution appears to be approximately Normal.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_dhistogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\AttributeTok{data =}\NormalTok{ dec\_means) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{gf\_dens}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{gf\_labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Histogram of 10,000 Bootstrapped Mean Dec Values"}\NormalTok{,}
          \AttributeTok{caption =} \StringTok{"Fig. 2: Bootstrap histogram of Dec values"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}

\hypertarget{standard-confidence-interval}{%
\subsection{Standard Confidence
Interval}\label{standard-confidence-interval}}

Confidence intervals are tools that are used to estimate a parameter.
Specifically, a confidence interval gives a range of values in which the
true value of the parameter may lie. An \(\alpha\)-level standard
confidence interval is given by
\[\hat{\theta}_S[\alpha] = \hat{\theta} \pm z_{\alpha}\hat{\sigma},\]
where \(\hat{\theta}\) is a point estimate of the parameter of interest
\(\theta\), \(\hat{\sigma}\) is the estimate of the standard deviation
of \(\hat{\theta}\) and \(z_{\alpha}\) is the \((100 *\alpha)\)th
percentile of the normal deviation \citep{Efron86}. We say that the
confidence interval constructed in this manner has a chance of capturing
the true parameter with a probability of \(\alpha\).

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. As such, the
standard confidence interval is sometimes called the normal confidence
interval for a bootstrapped sample. This means that for an unknown
skewed distribution, the standard confidence interval could present an
incorrect range. However, the same process can be used with bootstrap
sampling to form the bootstrap percentile method. This means that an
approximate bootstrap confidence interval will be created in the same
automatic way that the standard confidence interval was created. For
bootstrapped confidence intervals, the number of bootstrap replications
\(B\) must be large (around 2,000) due to the nature of confidence
intervals requiring greater accuracy \citep{Efron86}.

\hypertarget{standard-confidence-interval-in-use}{%
\subsubsection{Standard Confidence Interval in
Use}\label{standard-confidence-interval-in-use}}

Using our previous example to find a 95\% confidence interval for the
population mean of \texttt{Dec}, we would get the following confidence
interval:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FloatTok{0.025}\NormalTok{, }\FloatTok{0.975}\NormalTok{)) }\SpecialCharTok{*} 
  \FunctionTok{sd}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR)}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(SnowGR))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 13.85639 17.65957
\end{verbatim}

This means that we can say we are 95\% confident that the true mean
number of inches of snow that fall in December is betwwen 13.86 inches
and 17.66 inches. Our original histogram with the confidence interval is
below (Figure 3).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 3: Histogram with Standard Interval Shown"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{13.85639}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{17.65957}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}

\hypertarget{the-percentile-method}{%
\subsection{The Percentile Method}\label{the-percentile-method}}

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\([\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}]\)
\citep[\citet{EfronCasi}]{Efron86}. To go further, we can define
\(\hat{G}(t)\) as the bootstrap cdf, or the proportion of bootstrap
samples less than \(t\):
\[\hat{G}(t) = \frac{\#\{\hat{\theta}^{*b} \leq t\}}{B}.\] Thus the
\(\alpha\)th percentile point of the distribution is given by
\[\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).\]
It follows that the percentile interval can be represented as
\[\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].\] In the
case that the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is not usually the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] This
also can be represented as an interval,
\[\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ].\]
Therefore, the percentile method produces a correct interval for
\(\phi\) and due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist \citep{Efron86}.

\hypertarget{the-percentile-method-in-use}{%
\subsubsection{The Percentile Method in
Use}\label{the-percentile-method-in-use}}

Finding a 95\% confidence interval for the true population mean of
\texttt{Dec} using the percentile method, we get:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(}\FloatTok{0.025}\NormalTok{, }\FloatTok{0.975}\NormalTok{), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     2.5%    97.5% 
## 13.91170 17.72353
\end{verbatim}

Therefore, we can say that we are 95\% confident that the true
population mean of inches of snow that fall in December is between 13.91
inches and 17.72 inches. Our original histogram with the percentile
confidence interval is shown below (Figure 4).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 4: Histogram with Percentile Interval Shown"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{13.91170}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{17.72353}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf} This
interval is similar to the one created using the standard method, but
shifted slightly to the right.

\hypertarget{the-bias-corrected-bc-method}{%
\subsection{The Bias-Corrected (BC)
Method}\label{the-bias-corrected-bc-method}}

The next method we will be looking at is the bias-corrected percentile
method (BC method) which is an improvement upon the previous percentile
method because we now take into account the possibility of bias. It can
be shown that \(\hat{\theta}\) is biased upwards relative to \(\theta\)
which means that the confidence intervals should be adjusted downwards
\citep[\citet{EfronCasi}]{Efron86}. From our simulated bootstrap
replications
\(\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B},\)
define \[p_0 = \frac{\#\{\hat{\theta}^{*b} \leq \theta\}}{B},\] and
define the bias-correction value \[z_0 = \Phi^{-1}(p_0),\] where
\(\Phi^{-1}\) is the inverse function of the standard normal cdf. Thus,
we define a transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\), \[\hat{\phi} \sim N(\phi - z_0\tau, \tau^2),\] with \(z_0\)
and \(\tau\) constants. This means that we can say the bias-corrected
method has an \(\alpha\)-level endpoint which can be represented as
\[\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].\]
If \(\hat{G} = 0.50\), then half of the bootstrap distribution is less
than \(\hat{\theta}\) and our bias-correction value \(z_0 = 0\). In this
case, the confidence interval produced by BC would be the same interval
that was produced by the percentile method.

\hypertarget{the-bc-method-in-use}{%
\subsubsection{The BC Method in Use}\label{the-bc-method-in-use}}

Using this method to create a 95\% confidence interval has a couple more
steps because we need to find the bias-correction value. First we
calculate the sample mean:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_mean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{Dec, }\AttributeTok{data =}\NormalTok{ SnowGR); sample\_mean}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 15.75798
\end{verbatim}

Then we find the proportion of bootstrap replications that have a sample
mean less than the original sample mean.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{less\_than\_sample\_mean }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(dec\_means }\SpecialCharTok{\textless{}=}\NormalTok{ sample\_mean, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{))}\SpecialCharTok{/}\DecValTok{10000}
\NormalTok{less\_than\_sample\_mean}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.5208
\end{verbatim}

From this, we can calculate the bias-correction value, \(z_0\):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{z0 }\OtherTok{\textless{}{-}} \FunctionTok{qnorm}\NormalTok{(less\_than\_sample\_mean); z0}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.05216151
\end{verbatim}

Then we can find the modified percentiles and get the corresponding
confidence interval:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alphalow }\OtherTok{\textless{}{-}} \FloatTok{0.025}\NormalTok{; alphahigh }\OtherTok{\textless{}{-}} \FloatTok{0.975}

\NormalTok{newlow }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(}\DecValTok{2}\SpecialCharTok{*}\NormalTok{z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow)); }\CommentTok{\#newlow}
\NormalTok{newhigh }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(}\DecValTok{2}\SpecialCharTok{*}\NormalTok{z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh)); }\CommentTok{\#newhigh}

\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(newlow, newhigh), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 3.175238% 98.05047% 
##  14.01172  17.82017
\end{verbatim}

Therefore, we are 95\% confident that the true mean inches of snow that
fall in December in Grand Rapids, Michigan is between 14.01 inches and
17.82 inches. Our histogram (Figure 5) is shown below.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 5: Histogram with BC Interval Shown"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{14.01172}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{17.82017}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}

\hypertarget{the-bias-corrected-and-accelerated-bca-method}{%
\subsection{The Bias-Corrected and Accelerated (BCa)
Method}\label{the-bias-corrected-and-accelerated-bca-method}}

A further modification upon the BC interval is the bias-corrected and
accelerated method (BCa). For this method, we do not assume the the
standard error, \(\tau\), is constant as we did in the BC interval
\citep[\citet{EfronCasi}]{Efron86}. Rather, we assume the existence of a
monotone transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\),
\[\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2) \text{ where } \tau_\phi = 1+ a\phi.\]
The \(a\) is known as the acceleration and is a constant that describes
how the standard deviation of \(\hat{\phi}\) varies with \(\phi\). In
other words, \(a\) is proportional to the skewness of the bootstrap
distribution. For example, \(a=z_0\) for one-parameter exponential
families, however, there are many different algorithms to compute and
estimate \(a\) \citep{Flowers18}. Now, our \(\alpha\)-level endpoint for
a BCa confidence interval is
\[\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].\]
We can observe that if \(a = 0\), then
\(\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].\) When
calculating a BCa confidence interval, the acceleration value, \(a\), is
not a function of the bootstrap distribution and must be calculated
separately. However the process is algorithmic and can be calculated
without too much work. As we saw, each of the three previous methods
(percentile, BC, and BCa) all build upon each other and have less
restrictive assumptions, but this means that computation increases as we
loosen assumptions.

\hypertarget{the-bca-confidence-interval-in-use}{%
\subsubsection{The BCa Confidence Interval in
Use}\label{the-bca-confidence-interval-in-use}}

In practice, we can use the jackknife procedure to estimate the
acceleration value for this method. The jackknife procedure is a
different resampling technique that is often used to estimate the bias
and variance of a large population. Thus, we will run the jackknife
procedure:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{theta }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x)\{}\FunctionTok{mean}\NormalTok{(x)\}}
\NormalTok{jackknife\_results }\OtherTok{\textless{}{-}}\NormalTok{ bootstrap}\SpecialCharTok{::}\FunctionTok{jackknife}\NormalTok{(SnowGR}\SpecialCharTok{$}\NormalTok{Dec, theta) }
\end{Highlighting}
\end{Shaded}

We find the mean of the jackknife values below:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{jackmean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values); jackmean}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 15.75798
\end{verbatim}

Then we can compute an approximation for \(a\).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{estimated\_a }\OtherTok{\textless{}{-}}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{/}\DecValTok{6}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{sum}\NormalTok{((jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values }\SpecialCharTok{{-}}\NormalTok{ jackmean)}\SpecialCharTok{\^{}}\DecValTok{3}\NormalTok{)}\SpecialCharTok{/}
\NormalTok{  (}\FunctionTok{sum}\NormalTok{((jackknife\_results}\SpecialCharTok{$}\NormalTok{jack.values }\SpecialCharTok{{-}}\NormalTok{ jackmean)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))}\SpecialCharTok{\^{}}\NormalTok{(}\FloatTok{1.5}\NormalTok{); estimated\_a}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -0.02674911
\end{verbatim}

Now, we can find the adjusted percentiles and the bias-corrected with
acceleration confidence interval.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a\_newlow }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow))}\SpecialCharTok{/}
\NormalTok{                    (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow)))); }
\NormalTok{a\_newhigh }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh))}\SpecialCharTok{/}
\NormalTok{                     (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh)))); }
\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(a\_newlow, a\_newhigh), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 2.510119% 97.50908% 
##  13.91591  17.72353
\end{verbatim}

We are 95\% confident that the true mean number of inches of snow that
fall in December is between 13.92 inches and 17.72 inches. Figure 6
shows our confidence interval.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 6: Histogram with BCa Interval Shown"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{13.91591}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{17.72353}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1.pdf}

\hypertarget{the-bootstrap-t-studentized-method}{%
\subsection{The Bootstrap-t (Studentized)
Method}\label{the-bootstrap-t-studentized-method}}

The final method we explain is the bootstrap-t interval (or the
studentized method). Recall from earlier, our sample, \(X\) from which
we can calculate an estimate (\(\hat{\theta}(X)\)) of the parameter of
interest \(\theta\) \citep[\citet{Puth15}]{Efron86}. We can also
estimate \(\hat{\sigma}(X)\) for the standard error of \(\theta\). We
can use these parameters to find the Student's \(t\)-statistic defined
as: \[T = \frac{\hat{\theta} - \theta}{\hat{\sigma}}.\] As such, the
\(\alpha\)th percentile point of a confidence interval of \(\theta\)
would be \(\hat{\theta} - \hat{\sigma}T_{(\alpha)}\) where
\(T_{(\alpha)}\) represents the \(\alpha\)th percentile of the
\(t\)-distribution, \(T\). Unfortunately, the percentiles of this
\(t\)-distribution are unknown in most cases, but we can use
bootstrapping to estimate these percentiles. To do this, we perform a
large number, \(B\), of bootstrap samples, from which we can find the
bootstrap replications of the parameter of interest,
\(\hat{\theta^*} = \hat{\theta}(X^*)\), and the standard error,
\(\hat{\sigma^*} = \hat{\sigma}(X^*)\). From these we can calculate a
t-statistic for each bootstrap sample:
\[T^* = \frac{\hat{\theta}^* - \hat{\theta}}{\hat{\sigma}^*}.\] Using a
large number of these bootstrap samples, we can estimate the percentiles
of the \(t\)-distribution such that:
\[\hat{T}_{(\alpha)} = B*\alpha\text{th ordered value of all the bootstrap relications of }T^*.\]
This means that for \(B = 2,000\) and \(\alpha = 0.90\), then
\(\hat{T}_{(\alpha)}\) is the 1,800th ordered point of all of the
bootstrap replications of \(T^*\). It follows that the \(\alpha\)th
studentized confidence interval endpoint can be given with
\[\hat{\theta}_T[\alpha] = \hat{\theta} - \hat{\sigma}T_{(\alpha)},\]
where we can estimate the standard error using
\[\hat{\sigma} = \frac{1-\hat{\theta}^2}{\sqrt{n}}.\] One of the main
factors why the studentized method is so popular is because we assume
that our bootstrapped statistic is pivotal which means that the
confidence interval does not depend on any other parameters. Instead, we
can calculate the appropriate confidence interval for the parameter of
interest specifically from the bootstrapped statistics
\citep[\citet{Puth15}]{Efron86}.

\hypertarget{studentized-confidence-interval-in-use}{%
\subsubsection{Studentized Confidence Interval in
Use}\label{studentized-confidence-interval-in-use}}

Following the process above, we can use the following to get our
studentized confidence interval \citep{Lau20}:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}

\NormalTok{orig\_mean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\CommentTok{\# theta hat}
\NormalTok{orig\_se }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR)}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(SnowGR))}

\NormalTok{se }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nboot)}
\NormalTok{t\_stat }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nboot)}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{nboot) \{}
\NormalTok{  se[i] }\OtherTok{\textless{}{-}}\NormalTok{ sd[i]}\SpecialCharTok{/}\FunctionTok{sqrt}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(SnowGR))}
\NormalTok{  t\_stat[i] }\OtherTok{\textless{}{-}}\NormalTok{ (mean[i] }\SpecialCharTok{{-}}\NormalTok{ orig\_mean)}\SpecialCharTok{/}\NormalTok{se[i]}
\NormalTok{\}}

\NormalTok{dec\_t\_stat }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(t\_stat)}

\NormalTok{q }\OtherTok{\textless{}{-}} \FunctionTok{unname}\NormalTok{(}\FunctionTok{quantile}\NormalTok{(dec\_t\_stat}\SpecialCharTok{$}\NormalTok{t\_stat, }\FunctionTok{c}\NormalTok{(}\FloatTok{0.025}\NormalTok{, }\FloatTok{0.975}\NormalTok{)))}
\NormalTok{lower }\OtherTok{\textless{}{-}}\NormalTok{ q[}\DecValTok{1}\NormalTok{]}
\NormalTok{upper }\OtherTok{\textless{}{-}}\NormalTok{ q[}\DecValTok{2}\NormalTok{]}

\FunctionTok{c}\NormalTok{(orig\_mean }\SpecialCharTok{{-}}\NormalTok{ upper}\SpecialCharTok{*}\NormalTok{orig\_se, orig\_mean }\SpecialCharTok{{-}}\NormalTok{ lower}\SpecialCharTok{*}\NormalTok{orig\_se)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 14.03467 18.01709
\end{verbatim}

So we are 95\% confident that the true mean number of inches of snow
that fall in December in Grand Rapids, Michigan is between 14.03 inches
and 18.02 inches. This would give us the following confidence interval
(Figure 7)

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{, }
       \AttributeTok{caption =} \StringTok{"Fig. 7: Histogram with Studentized Interval Shown"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{14.03467}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept=} \FloatTok{18.01709}\NormalTok{)) }
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-17-1.pdf}

\hypertarget{simulation}{%
\section{Simulation}\label{simulation}}

For our simulation, we will randomly sample from distributions in which
we know the true population mean. For example, we will be sampling from
a Gamma distribution where we can find the true population mean using
the shape and rate parameters of the sample distribution. After we
sample from the distribution, we will create a bootstrapped confidence
interval for a few of the methods we described in the exposition. Since
we know the true population mean of the distribution, we will determine
if the true population mean is contained in the confidence interval for
each of the methods we are testing. We will perform this process a large
number of times (10,000 simulations) and then average by the number of
iterations to determines on average, the proportion that each
bootstrapped confidence interval method contains the true population
mean for a given distribution. We will do this for a couple different
distributions as well as a range of sample sizes to show how the
intervals perform with skewed distributions and larger sample sizes.

To do this simulation, we will be using the \texttt{boot} package
\citep{boot}. This package can compute a bootstrap object which contains
the output from a bootstrap calculation of a given sample and then
produces confidence intervals from the bootstrap object. In particular,
we will use the \texttt{boot.ci} function to create four types of
confidence intervals: the standard (or normal) confidence interval, the
studentized confidence interval, the percentile confidence interval, and
the bias-corrected with acceleration confidence interval.

To start, we will declare some constants including the sample sizes and
the number of bootstrap replications we are using.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_sizes }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{60}\NormalTok{, }\DecValTok{100}\NormalTok{, }\DecValTok{150}\NormalTok{)}
\NormalTok{num\_sample\_sizes }\OtherTok{\textless{}{-}} \FunctionTok{length}\NormalTok{(sample\_sizes)}
\NormalTok{boot\_reps }\OtherTok{\textless{}{-}} \DecValTok{2000}
\end{Highlighting}
\end{Shaded}

Next, we create a simulation that randomly selects a sample from a
Normal distribution and performs a bootstrap with 2,000 bootstrap
replications. Then, we create the confidence intervals from this
bootstrap and determine if the true mean, our population parameter, is
contained within our intervals. Since we are using a \(Normal(0,1)\)
distribution, our true population mean is equal to 0. We will repeat
this process 10,000 times as specified for each of the sample sizes that
we are testing. Since we are originally sampling from a Normal
distribution, the assumptions we hold for the standard interval are true
and thus, we would expect all of the methods to perform comparably to
each other at large sample sizes.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{run\_sim\_normal }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}

\NormalTok{  boot.stat }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

\NormalTok{  master }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\AttributeTok{nrow =}\NormalTok{ num\_sample\_sizes, }\AttributeTok{ncol =} \DecValTok{5}\NormalTok{)}

  \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{num\_sample\_sizes) \{}

\NormalTok{    norm\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    student\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    percentile\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    bca\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}

\NormalTok{    n }\OtherTok{\textless{}{-}}\NormalTok{ sample\_sizes[i]}
\NormalTok{    true\_mean }\OtherTok{\textless{}{-}} \DecValTok{0}

    \ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{sim\_reps) \{}
\NormalTok{      X }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(n, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}
\NormalTok{      values }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(X)}
\NormalTok{      results }\OtherTok{\textless{}{-}} \FunctionTok{boot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ values, }\AttributeTok{statistic =}\NormalTok{ boot.stat, }\AttributeTok{R =}\NormalTok{ boot\_reps)}
\NormalTok{      ci }\OtherTok{\textless{}{-}} \FunctionTok{boot.ci}\NormalTok{(results, }\AttributeTok{conf =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{type =} \FunctionTok{c}\NormalTok{(}\StringTok{"norm"}\NormalTok{, }\StringTok{"perc"}\NormalTok{, }\StringTok{"bca"}\NormalTok{, }\StringTok{"stud"}\NormalTok{))}

\NormalTok{      norm\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{normal[}\DecValTok{2}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{normal[}\DecValTok{3}\NormalTok{]}
\NormalTok{      student\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{student[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{student[}\DecValTok{5}\NormalTok{]}
\NormalTok{      percentile\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{percent[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{percent[}\DecValTok{5}\NormalTok{]}
\NormalTok{      bca\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{bca[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{bca[}\DecValTok{5}\NormalTok{]}
\NormalTok{    \}}

\NormalTok{    master[i, }\DecValTok{1}\NormalTok{] }\OtherTok{\textless{}{-}}\NormalTok{ sample\_sizes[i]}
\NormalTok{    master[i, }\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(norm\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{3}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(student\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{4}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(percentile\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{5}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(bca\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{  \}}

\NormalTok{  master\_df }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(master) }\SpecialCharTok{\%\textgreater{}\%} 
      \FunctionTok{rename}\NormalTok{(}\StringTok{"Sample\_Size"} \OtherTok{=}\NormalTok{ V1, }\StringTok{"Normal"} \OtherTok{=}\NormalTok{ V2, }\StringTok{"Studentized"} \OtherTok{=}\NormalTok{ V3, }
             \StringTok{"Percentile"} \OtherTok{=}\NormalTok{ V4, }\StringTok{"BCa"} \OtherTok{=}\NormalTok{ V5) }
\NormalTok{  master\_df}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Secondly, we select from a Gamma Distribution with a shape parameter of
1 and a scale parameter of 4 (\(Gamma(1,4)\)). Therefore, the true
population mean is 4 because the true mean is given by the product of
the shape and scale parameter. This simulation follows the same process
as above where we sampled from the Normal distribution except we are
sampling from a Gamma distribution this time. Since, our Gamma
distribution is heavily left-skewed, we would expect the studentized
confidence interval and the BCa confidence interval to perform the best.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{run\_sim\_gamma }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}

\NormalTok{  boot.stat }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

\NormalTok{  master }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\AttributeTok{nrow =}\NormalTok{ num\_sample\_sizes, }\AttributeTok{ncol =} \DecValTok{5}\NormalTok{)}

  \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{num\_sample\_sizes) \{}

\NormalTok{    norm\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    student\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    percentile\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    bca\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}

\NormalTok{    n }\OtherTok{\textless{}{-}}\NormalTok{ sample\_sizes[i]}
\NormalTok{    true\_mean }\OtherTok{\textless{}{-}} \DecValTok{4}

    \ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{sim\_reps) \{}
\NormalTok{      X }\OtherTok{\textless{}{-}} \FunctionTok{rgamma}\NormalTok{(sample\_sizes, }\AttributeTok{shape =} \DecValTok{1}\NormalTok{, }\AttributeTok{scale =} \DecValTok{4}\NormalTok{)}
\NormalTok{      values }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(X)}

\NormalTok{      results }\OtherTok{\textless{}{-}} \FunctionTok{boot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ values, }\AttributeTok{statistic =}\NormalTok{ boot.stat, }\AttributeTok{R =}\NormalTok{ boot\_reps)}
\NormalTok{      ci }\OtherTok{\textless{}{-}} \FunctionTok{boot.ci}\NormalTok{(results, }\AttributeTok{conf =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{type =} \FunctionTok{c}\NormalTok{(}\StringTok{"norm"}\NormalTok{, }\StringTok{"perc"}\NormalTok{, }\StringTok{"bca"}\NormalTok{, }\StringTok{"stud"}\NormalTok{))}

\NormalTok{      norm\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{normal[}\DecValTok{2}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{normal[}\DecValTok{3}\NormalTok{]}
\NormalTok{      student\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{student[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{student[}\DecValTok{5}\NormalTok{]}
\NormalTok{      percentile\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{percent[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{percent[}\DecValTok{5}\NormalTok{]}
\NormalTok{      bca\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{bca[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{bca[}\DecValTok{5}\NormalTok{]}
\NormalTok{    \}}
\NormalTok{    master[i, }\DecValTok{1}\NormalTok{] }\OtherTok{\textless{}{-}}\NormalTok{ sample\_sizes[i]}
\NormalTok{    master[i, }\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(norm\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{3}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(student\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{4}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(percentile\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{5}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(bca\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{  \}}
\NormalTok{  master\_df }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(master) }\SpecialCharTok{\%\textgreater{}\%} 
      \FunctionTok{rename}\NormalTok{(}\StringTok{"Sample\_Size"} \OtherTok{=}\NormalTok{ V1, }\StringTok{"Normal"} \OtherTok{=}\NormalTok{ V2, }\StringTok{"Studentized"} \OtherTok{=}\NormalTok{ V3, }
             \StringTok{"Percentile"} \OtherTok{=}\NormalTok{ V4, }\StringTok{"BCa"} \OtherTok{=}\NormalTok{ V5) }
\NormalTok{  master\_df}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Lastly, we will be sampling from a Log-Normal distribution. This means
that if we have a random variable, \(X \sim Lognormal\), then the random
variable, \(Y = ln(X) \sim Normal\). Therefore, the distribution is
left-skewed, although not as heavily as the Gamma distribution we
previously used. The true mean of this distribution can be found using
\(E(X) = e^{u+\frac{\sigma^2}{2}}\) which means that for a
\(Lognormal(0,1)\) distribution, we expect the population mean to be
\(e^{\frac{1}{2}}\). Again, we would expect the studentized confidence
interval and the BCa confidence interval to perform the best.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{run\_sim\_log\_normal }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}

\NormalTok{  boot.stat }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

\NormalTok{  master }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\AttributeTok{nrow =}\NormalTok{ num\_sample\_sizes, }\AttributeTok{ncol =} \DecValTok{5}\NormalTok{)}

  \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{num\_sample\_sizes) \{}

\NormalTok{    norm\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    student\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    percentile\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}
\NormalTok{    bca\_contains }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, sim\_reps)}

\NormalTok{    n }\OtherTok{\textless{}{-}}\NormalTok{ sample\_sizes[i]}
\NormalTok{    true\_mean }\OtherTok{\textless{}{-}} \FunctionTok{exp}\NormalTok{(}\DecValTok{1}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}

    \ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{sim\_reps) \{}
\NormalTok{      X }\OtherTok{\textless{}{-}} \FunctionTok{rlnorm}\NormalTok{(n, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)}
\NormalTok{      values }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(X)}

\NormalTok{      results }\OtherTok{\textless{}{-}} \FunctionTok{boot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ values, }\AttributeTok{statistic =}\NormalTok{ boot.stat, }\AttributeTok{R =}\NormalTok{ boot\_reps)}
\NormalTok{      ci }\OtherTok{\textless{}{-}} \FunctionTok{boot.ci}\NormalTok{(results, }\AttributeTok{conf =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{type =} \FunctionTok{c}\NormalTok{(}\StringTok{"norm"}\NormalTok{, }\StringTok{"perc"}\NormalTok{, }\StringTok{"bca"}\NormalTok{, }\StringTok{"stud"}\NormalTok{))}

\NormalTok{      norm\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{normal[}\DecValTok{2}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{normal[}\DecValTok{3}\NormalTok{]}
\NormalTok{      student\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{student[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{student[}\DecValTok{5}\NormalTok{]}
\NormalTok{      percentile\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{percent[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{percent[}\DecValTok{5}\NormalTok{]}
\NormalTok{      bca\_contains[j] }\OtherTok{\textless{}{-}}\NormalTok{ true\_mean }\SpecialCharTok{\textgreater{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{bca[}\DecValTok{4}\NormalTok{] }\SpecialCharTok{\&}\NormalTok{ true\_mean }\SpecialCharTok{\textless{}=}\NormalTok{ ci}\SpecialCharTok{$}\NormalTok{bca[}\DecValTok{5}\NormalTok{]}
\NormalTok{    \}}
\NormalTok{    master[i, }\DecValTok{1}\NormalTok{] }\OtherTok{\textless{}{-}}\NormalTok{ sample\_sizes[i]}
\NormalTok{    master[i, }\DecValTok{2}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(norm\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{3}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(student\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{4}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(percentile\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{    master[i, }\DecValTok{5}\NormalTok{] }\OtherTok{\textless{}{-}} \FunctionTok{sum}\NormalTok{(bca\_contains)}\SpecialCharTok{/}\NormalTok{sim\_reps}
\NormalTok{  \}}
\NormalTok{  master\_df }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(master) }\SpecialCharTok{\%\textgreater{}\%} 
      \FunctionTok{rename}\NormalTok{(}\StringTok{"Sample\_Size"} \OtherTok{=}\NormalTok{ V1, }\StringTok{"Normal"} \OtherTok{=}\NormalTok{ V2, }\StringTok{"Studentized"} \OtherTok{=}\NormalTok{ V3, }
             \StringTok{"Percentile"} \OtherTok{=}\NormalTok{ V4, }\StringTok{"BCa"} \OtherTok{=}\NormalTok{ V5) }
\NormalTok{  master\_df}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Once we have all the simulations, we can run them below.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{normal\_df\_trial2 }\OtherTok{\textless{}{-}} \FunctionTok{run\_sim\_normal}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
\FunctionTok{write.csv}\NormalTok{(normal\_df\_trial2,}\StringTok{"normal\_data\_df\_trial2.csv"}\NormalTok{, }\AttributeTok{row.names =} \ConstantTok{FALSE}\NormalTok{)}

\NormalTok{log\_normal\_df\_trial2 }\OtherTok{\textless{}{-}} \FunctionTok{run\_sim\_log\_normal}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
\FunctionTok{write.csv}\NormalTok{(log\_normal\_df\_trial2,}\StringTok{"log\_normal\_df\_trial2.csv"}\NormalTok{, }\AttributeTok{row.names =} \ConstantTok{FALSE}\NormalTok{)}

\NormalTok{gamma\_df\_trial2 }\OtherTok{\textless{}{-}} \FunctionTok{run\_sim\_gamma}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
\FunctionTok{write.csv}\NormalTok{(gamma\_df\_trial2,}\StringTok{"gamma\_data\_df\_trial2.csv"}\NormalTok{, }\AttributeTok{row.names =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{normal\_df }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\AttributeTok{file =} \StringTok{\textquotesingle{}normal\_data\_df\_trial2.csv\textquotesingle{}}\NormalTok{)}
\NormalTok{gamma\_df }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\AttributeTok{file =} \StringTok{\textquotesingle{}gamma\_data\_df\_trial2.csv\textquotesingle{}}\NormalTok{)}
\NormalTok{log\_normal\_df }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\AttributeTok{file =} \StringTok{\textquotesingle{}log\_normal\_df\_trial2.csv\textquotesingle{}}\NormalTok{)}

\CommentTok{\# create the figure that shows the normal output }
\NormalTok{normal\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

\NormalTok{normal\_fig }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ normal\_df\_long, }
                    \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Sample\_Size, }\AttributeTok{y =}\NormalTok{ prop, }\AttributeTok{color =}\NormalTok{ method)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Proportion of CIs that Contain the True Parameter from Normal }
\StringTok{       Distibution"}\NormalTok{, }\AttributeTok{x =} \StringTok{"Sample Size"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Proportion (of 10,000 Simulations)"}\NormalTok{, }\AttributeTok{color =} \StringTok{"Method"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for}
\StringTok{       different sample sizes on Normal Dist."}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_light}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position=}\StringTok{"bottom"}\NormalTok{)}

\CommentTok{\# create the table for the normal output}
\NormalTok{normal\_table }\OtherTok{\textless{}{-}}\NormalTok{ normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled from Normal Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(normal\_df)))}

\CommentTok{\# create the figure that shows the gamma output}
\NormalTok{gamma\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ gamma\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

\NormalTok{gamma\_fig }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ gamma\_df\_long, }
                    \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Sample\_Size, }\AttributeTok{y =}\NormalTok{ prop, }\AttributeTok{color =}\NormalTok{ method)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Proportion of CIs that Contain the True Parameter from Gamma }
\StringTok{       Distibution"}\NormalTok{, }\AttributeTok{x =} \StringTok{"Sample Size"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Proportion (of 10,000 Simulations)"}\NormalTok{, }\AttributeTok{color =} \StringTok{"Method"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for}
\StringTok{       different sample sizes on Gamma Dist."}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_light}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position=}\StringTok{"bottom"}\NormalTok{)}

\CommentTok{\# create the table for the gamma output}
\NormalTok{gamma\_table }\OtherTok{\textless{}{-}}\NormalTok{ gamma\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled From Gamma Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(gamma\_df)), }\AttributeTok{digits =} \DecValTok{3}\NormalTok{) }

\CommentTok{\# create the figure that shows the log normal output}
\NormalTok{log\_normal\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ log\_normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{5}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

\NormalTok{log\_normal\_fig }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ log\_normal\_df\_long, }
                    \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Sample\_Size, }\AttributeTok{y =}\NormalTok{ prop, }\AttributeTok{color =}\NormalTok{ method)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Proportion of CIs that Contain the True Parameter from Log{-}Normal }
\StringTok{       Distibution"}\NormalTok{, }\AttributeTok{x =} \StringTok{"Sample Size"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Proportion (of 10,000 Simulations)"}\NormalTok{, }\AttributeTok{color =} \StringTok{"Method"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for}
\StringTok{       different sample sizes on Log{-}Normal Dist."}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_light}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position=}\StringTok{"bottom"}\NormalTok{)}

\CommentTok{\# create the table for the log normal output}
\NormalTok{log\_normal\_table }\OtherTok{\textless{}{-}}\NormalTok{ log\_normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled from Normal Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(log\_normal\_df)))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log\_normal\_fig}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/output the graphs/tables-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log\_normal\_table}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:create graphs}Sampled from Normal Dist.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
Sample Size & Normal & Basic & Studentized & Percentile & BCa\\
\midrule
4 & 0.6795 & 0.6542 & 0.8186 & 0.6810 & 0.6814\\
10 & 0.7911 & 0.7684 & 0.8526 & 0.8019 & 0.8266\\
15 & 0.8268 & 0.8061 & 0.8705 & 0.8387 & 0.8552\\
20 & 0.8437 & 0.8236 & 0.8797 & 0.8526 & 0.8725\\
30 & 0.8639 & 0.8464 & 0.8951 & 0.8734 & 0.8900\\
\addlinespace
40 & 0.8789 & 0.8623 & 0.9055 & 0.8881 & 0.9013\\
60 & 0.8897 & 0.8783 & 0.9100 & 0.8978 & 0.9069\\
80 & 0.9088 & 0.8969 & 0.9218 & 0.9136 & 0.9181\\
100 & 0.9103 & 0.8980 & 0.9264 & 0.9162 & 0.9229\\
\bottomrule
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{normal\_fig}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/output the graphs/tables-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{normal\_table}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:create graphs}Sampled from Normal Dist.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
Sample Size & Normal & Basic & Studentized & Percentile & BCa\\
\midrule
4 & 0.8046 & 0.7913 & 0.9058 & 0.7941 & 0.7943\\
10 & 0.9015 & 0.9005 & 0.9265 & 0.9007 & 0.8994\\
15 & 0.9187 & 0.9183 & 0.9331 & 0.9184 & 0.9167\\
20 & 0.9257 & 0.9249 & 0.9357 & 0.9242 & 0.9235\\
30 & 0.9363 & 0.9372 & 0.9421 & 0.9351 & 0.9323\\
\addlinespace
40 & 0.9421 & 0.9421 & 0.9466 & 0.9416 & 0.9416\\
60 & 0.9381 & 0.9390 & 0.9420 & 0.9389 & 0.9386\\
80 & 0.9403 & 0.9408 & 0.9430 & 0.9406 & 0.9396\\
100 & 0.9483 & 0.9478 & 0.9499 & 0.9475 & 0.9480\\
\bottomrule
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gamma\_fig}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/output the graphs/tables-3.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gamma\_table}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:create graphs}Sampled From Gamma Dist.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
Sample Size & Normal & Basic & Studentized & Percentile & BCa\\
\midrule
4 & 0.849 & 0.833 & 0.904 & 0.858 & 0.875\\
10 & 0.848 & 0.829 & 0.901 & 0.854 & 0.871\\
15 & 0.851 & 0.832 & 0.901 & 0.856 & 0.871\\
20 & 0.846 & 0.827 & 0.903 & 0.851 & 0.867\\
30 & 0.838 & 0.822 & 0.895 & 0.846 & 0.864\\
\addlinespace
40 & 0.848 & 0.831 & 0.899 & 0.853 & 0.866\\
60 & 0.847 & 0.829 & 0.907 & 0.857 & 0.874\\
80 & 0.849 & 0.831 & 0.902 & 0.854 & 0.872\\
100 & 0.848 & 0.831 & 0.897 & 0.853 & 0.866\\
\bottomrule
\end{tabular}
\end{table}

\bibliographystyle{agsm}
\bibliography{bibliography.bib}

\end{document}